\chapter{Personal review}
\label{chap:review}

My interest for computers and programming has led me to choose this topic for my work and it is not a choice I regret.

QR-Codes were a subject I personally wanted to understand for some time but never got round to it. Having the possibility to use my knowledge in Python to apply what I had researched was also very satisfying and self-rewarding. I hope I succeeded in trying to explain the inner workings and principles making these great inventions possible. I appreciated the great freedom both on the subject's choice and the realization. Apart from learning about barcodes and QR-Codes, I have also had the opportunity to put to use my English language skills.
Using \LaTeX\  helped practising problem solving as many things can be done in multiple ways and I needed to find the best or easiest method, especially regarding tables.

The Python scripts also represent a great part of this work. While not perfect implementations, they offer a better understanding of the sometimes abstract concepts discussed and allow the generation steps to be visualized personally, using any value.

Overall, I am pleased with the fruit of my research and understanding. My only regret would be that I have not yet been able to fully understand the intricacies of Reed-Solomon codes, but I hope that one day I will.

Finally, I would like to thank Mr. Erspamer for his availability, his support and his insightful feedbacks. I must also credit my brother, who allowed me to obtain some documents such as the ISO standards\cite{ISO16388}\cite{ISO18004} and who helped with the printing of my work.
